`r paste(params$species)` biomass trends and figures

### 2019-12-11 analysis/VOCC/3-bioclimatic-trends.Rmd
move bioclim functions to package; add option for tv-depth model; add 2009-2011 year pair

---
title: "`r paste(params$species)` biomass trends and figures"
author: "Philina English"
date: "`r Sys.Date()`"
output: html_document
params: 
    #species: "Pacific Cod" # 
    species: "Redbanded Rockfish"
    region: "All synoptic surveys"
    start_year: 2008
    bind_predictions: FALSE
    covs: "-fixed"  
    # year_cutoff: 2012
    #covs: "-log-both-fixed-depth"
    #covs: "-log-both-noAR1" # covariate string used to name model rds 
---

@@

  adult_predictions_obj <- predict(adult_biomass, newdata = nd, return_tmb_object = TRUE)
  adult_predictions <- adult_predictions_obj$data

  # write as function
  bioclim_by_year <- function(yr, model_for_prior_year, pred_grid) {
    nd <- pred_grid

    nd_trim <- nd %>% filter(year < yr)
    fake_years <- nd %>%
      mutate(year = year - 2) %>%
      filter(year > 2007) %>%
      filter(year < yr)

    fake <- sort(unique(fake_years$year))
    true <- sort(unique(nd$year))

    if (!identical(fake, true)) "Something is wrong with the pred_grid years."

    prior_years <- predict(model_for_prior_year, newdata = nd_trim, return_tmb_object = FALSE)
    fake_est <- predict(model_for_prior_year, newdata = fake_years, return_tmb_object = FALSE)
    fake_est$year <- fake_est$year + 2

    bioclim_start <- filter(prior_years, year >= yr - 2) %>%
      filter(year <= yr - 1) %>%
      mutate(bioclim = est)
    bioclim_end <- filter(fake_est, year >= yr) %>%
      filter(year <= yr + 1) %>%
      mutate(bioclim = est_non_rf + omega_s)
    bioclim_all <- dplyr::bind_rows(bioclim_start, bioclim_end) %>%
      select(X, Y, year, bioclim)

    bioclim_all
  }

@@

  # for multiyear bioclim
  bioclim_multi_year <- function(yr, model_for_prior_year, pred_grid) {
    nd <- pred_grid
    period2_range <- max(nd$year) - yr + 1

    nd_trim <- nd %>% filter(year < yr)
    fake_years <- nd %>%
      mutate(year = year - 2) %>%
      filter(year > 2007) %>%
      filter(year < yr)

    fake <- sort(unique(fake_years$year))
    true <- sort(unique(model_for_prior_year$data$year))

    if (!identical(fake, true)) "Model years doesn't match filtered pred_grid years."

    prior_years <- predict(model_for_prior_year, newdata = nd_trim, return_tmb_object = FALSE)

    past_predictions <- prior_years %>%
      mutate(past_est = est, bioclim = est) %>%
      select(X, Y, year, past_est, bioclim)

    fake_est <- predict(model_for_prior_year, newdata = fake_years, return_tmb_object = FALSE)
    fake_est$year <- fake_est$year + 2

    bioclim <- filter(fake_est, year >= yr + 2 - 2) %>%
      filter(year <= yr + 2 - 1) %>%
      mutate(bioclim = est_non_rf + omega_s)

    if (period2_range >= 4) {
      fake_minus4 <- nd %>%
        mutate(year = year - 4) %>%
        filter(year > 2007)

      fake <- sort(unique(fake_minus4$year))
      true <- sort(unique(nd_trim$year))

      if (!identical(fake, true)) "Model years doesn't match filtered pred_grid years -4."

      fake_est_minus4 <- predict(model_for_prior_year, newdata = fake_minus4, return_tmb_object = FALSE)
      fake_est_minus4$year <- fake_est_minus4$year + 4
      # sort(unique(fake_est_minus4$year))

      bioclim_4 <- filter(fake_est_minus4, year >= yr + 4 - 2) %>%
        filter(year <= yr + 4 - 1) %>%
        mutate(bioclim = est_non_rf + omega_s)

      bioclim <- dplyr::bind_rows(bioclim, bioclim_4)
    }

    if (period2_range >= 6) {
      fake_minus6 <- nd %>%
        mutate(year = year - 6) %>%
        filter(year > 2007)
      fake <- sort(unique(fake_minus6$year))
      true <- sort(unique(nd_trim$year))

      if (!identical(fake, true)) "Model years doesn't match filtered pred_grid years -6."

      fake_est_minus6 <- predict(model_for_prior_year, newdata = fake_minus6, return_tmb_object = FALSE)
      fake_est_minus6$year <- fake_est_minus6$year + 6

      bioclim_6 <- filter(fake_est_minus6, year >= yr + 6 - 2) %>%
        filter(year <= yr + 6 - 1) %>%
        mutate(bioclim = est_non_rf + omega_s)

      bioclim <- dplyr::bind_rows(bioclim, bioclim_6)
    }

    bioclim_all <- dplyr::bind_rows(past_predictions, bioclim) %>%
      select(X, Y, year, past_est, bioclim)

    bioclim_all
  }

  model_pre_2015 <- readRDS(paste0(
    "data/", spp,
    "/mod-mat-biomass-", spp, covs, "-past-2015-", ssid_string, "-prior-", priors, ".rds"
  ))

  adult_predictions <- left_join(adult_predictions, bioclim_blob)
  adult_predictions <- left_join(adult_predictions, bioclim_2013)
  adult_predictions <- left_join(adult_predictions, bioclim_2015)
  adult_predictions <- left_join(adult_predictions, bioclim_2017)

  View(adult_predictions)

@@

  saveRDS(imm_predictions, file = paste0(
    "data/", spp,
    "/bioclimatic-predictions-", spp, covs, "-", ssid_string, "-imm-2015-cutoff.rds"
  ))
} else { # need to add non-split-by-age version
  # total_biomass <- readRDS(paste0("data/", spp,
  #   "/model-total-biomass-", spp, "-fixed", covs, "-", ssid_string, "-prior-", priors, ".rds"))
  #
  # nd_all <- readRDS(paste0("data/predicted-DO-new.rds"))
  # nd <- nd_all %>%
  #   filter(ssid %in% model_ssid) %>%
  #   filter(year %in% unique(total_biomass$data$year))%>%
  # mutate(do_mlpl_scaled = (log_do - total_biomass$data$do_mlpl_mean) /(total_biomass$data$do_mlpl_sd*2), do_mlpl_scaled2 = do_mlpl_scaled^2)
  # nd <- na.omit(nd)
  # nd$year <- as.integer(nd$year)
  #
  # predictions <- predict(total_biomass, newdata = nd)
  # saveRDS(predictions, file = paste0("data/", spp,
  #   "/predictions-", spp, covs, "-", ssid_string, "-total-bioclimatic-prior-", priors, ".rds"))
}
```

@@

  p_adult_st <- list()
  for (i in seq_len(length(model_ssid))) {
    p_adult_st[[i]] <- plot_facet_map(ad_prediction_list[[i]], "epsilon_st",
      white_zero = TRUE,
      high_fill = "#66C2A5",
      low_fill = "#5E4FA2"
      # raster_limits = c(0, max_raster),
      # legend_position = "none",
      # transform_col = fourth_root_power
    ) +
      # labs(fill = "kg/ha") +
      ggtitle(paste0("", species, " mature spatiotemporal variation"))
    p_adult_st[[i]]
  }
  print(p_adult_st)
  # } else {
  # p_adult_st <- plot_facet_map(predictions, "epsilon_st",
  #   transform_col = fourth_root_power
  # ) +
  #   #labs(fill = "kg/ha") +
  #   ggtitle(paste0("", species, "total spatiotemporal variation"))
  #
  # print(p_adult_st)
}
```

>>>>

---
title: "`r paste(params$species)` biomass trends and figures"
author: "Philina English"
date: "`r Sys.Date()`"
output: html_document
params: 
    # species: "Redbanded Rockfish"
    # species: "Pacific Cod" 
    species: "Pacific Ocean Perch" 
    region: "All synoptic surveys"
    start_year: 2008
    bind_predictions: FALSE
    covs: "-fixed"  
    # covs: "-tv-depth"
    # year_cutoff: 2012
    #covs: "-log-both-fixed-depth"
    #covs: "-log-both-noAR1" # covariate string used to name model rds 
---

@@

  adult_predictions_obj <- predict(adult_biomass, newdata = nd, return_tmb_object = TRUE)

  # multiyear bioclimatic calculation
  model_pre_2015 <- readRDS(paste0(
    "data/", spp,
    "/mod-mat-biomass-", spp, covs, "-past-2015-", ssid_string, "-prior-", priors, ".rds"
  ))

  adult_predictions <- adult_predictions_obj$data
  adult_predictions <- left_join(adult_predictions, bioclim_blob)

  # saveRDS(adult_predictions, file = paste0(
  #   "data/", spp,
  #   "/bioclimatic-predictions-", spp, covs, "-", ssid_string, "-mat-2015-cutoff.rds"
  # ))
  # 
  # add biennial bioclimatic calculation
  model_pre_2011 <- readRDS(paste0(
    "data/", spp,
    "/mod-mat-biomass-", spp, covs, "-past-2011-", ssid_string, "-prior-", priors, ".rds"
  ))
  sort(unique(model_pre_2011$data$year))

  bioclim_2011 <- bioclim_by_year(2011, model_pre_2011, nd)
  bioclim_2011 <- rename(bioclim_2011, bioclim_2011 = bioclim)
  glimpse(bioclim_2011)

@@

  adult_predictions <- left_join(adult_predictions, bioclim_2011)
  adult_predictions <- left_join(adult_predictions, bioclim_2013)
  adult_predictions <- left_join(adult_predictions, bioclim_2015)
  adult_predictions <- left_join(adult_predictions, bioclim_2017)

#  View(adult_predictions)

@@

  saveRDS(imm_predictions, file = paste0(
    "data/", spp,
    "/bioclimatic-predictions-", spp, covs, "-", ssid_string, "-imm-2015-cutoff.rds"
  ))

} else { # need to add non-split-by-age version
  # total_biomass <- readRDS(paste0("data/", spp,
  #   "/model-total-biomass-", spp, "-fixed", covs, "-", ssid_string, "-prior-", priors, ".rds"))
  #
  # nd_all <- readRDS(paste0("data/predicted-DO-new.rds"))
  # nd <- nd_all %>%
  #   filter(ssid %in% model_ssid) %>%
  #   filter(year %in% unique(total_biomass$data$year))%>%
  # mutate(do_mlpl_scaled = (log_do - total_biomass$data$do_mlpl_mean) /(total_biomass$data$do_mlpl_sd*2), do_mlpl_scaled2 = do_mlpl_scaled^2)
  # nd <- na.omit(nd)
  # nd$year <- as.integer(nd$year)
  #
  # predictions <- predict(total_biomass, newdata = nd)
  # saveRDS(predictions, file = paste0("data/", spp,
  #   "/predictions-", spp, covs, "-", ssid_string, "-total-bioclimatic-prior-", priors, ".rds"))
}
```

@@

  p_adult_st <- list()
  for (i in seq_len(length(model_ssid))) {
    p_adult_st[[i]] <- plot_facet_map(ad_prediction_list[[i]], "do_est",
      # white_zero = TRUE,
      # high_fill = "#66C2A5",
      # low_fill = "#5E4FA2"
      # # raster_limits = c(0, max_raster),
      # legend_position = "none",
      # transform_col = fourth_root_power
    ) +
      # labs(fill = "kg/ha") +
      ggtitle(paste0("", species, " mature spatiotemporal variation"))
    p_adult_st[[i]]
  }
  print(p_adult_st)
  # } else {
  # p_adult_st <- plot_facet_map(predictions, "epsilon_st",
  #   transform_col = fourth_root_power
  # ) +
  #   #labs(fill = "kg/ha") +
  #   ggtitle(paste0("", species, "total spatiotemporal variation"))
  #
  # print(p_adult_st)
}
```
